Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2024-05-30 SS3_Krill_481.Rmd
commit 31 may

---
title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
author:
  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
  - César Cárdenas^[Instituto Antártico Chileno, Chile]
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others. Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations. This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques. All this data was proceses in @Dornam2021; 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics in the Western Antarctic Peninsula (WAP) area.  Data processing involved sophisticated techniques, including data cleaning, spatial manipulation, rasterization, and projection transformations. These processes were facilitated by the R libraries *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], ensuring high-quality and accurate environmental datasets. To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020]. All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator **"fleet"** to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.

@@

\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population.

This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.

@@

Like most invertebrates, krill is a species for which aging is complex. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Despite this, integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.

@@

Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics. Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds. Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem. By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.

@@

The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill. By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model. This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.

In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3     | "s2" Without S-R relation         |
|    s4     | "s2" Ricker S-R relation estimated          |
|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |

@@

## Kobe (status)

@@

@Kinzey2015a, @Kinzey2011 and @Kinzey2019a used integrated models with fishery and survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1. Additionally, @Wang2021 made significant contributions to integrated statistical models for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings. Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) and El Niño.

Wang..
Lo que es nuevo aca es

**Paragraph 3:**  
From an ecological standpoint, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean. In terms of the krill fishery, this area includes a significant fishing ground, specifically subarea 48.1. Notably, recent years have seen a higher concentration of krill catches in this subarea, both temporally and spatially.

@@

>>>>

---
title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
author:
  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
  - César Cárdenas^[Instituto Antártico Chileno, Chile]
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others. Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations. This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques. All this data was proceses in @Dornam2021; 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics in the Western Antarctic Peninsula (WAP) area.  To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020]. All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator **"fleet"** to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.

@@

\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population. This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.

Life history arameters taken as priors taken from @Smith2023a and @Maschette2020

@@

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Despite this, integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.

@@

Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.  In this analysis we incorporated penguin population index as predator *"fleets"*. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock].

@@

To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model  regardng this configurations. In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3     | "s2" without S-R relation         |
|    s4     | "s2" Ricker S-R relation estimated          |
|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |

@@

models S01, S1, S2, and S5 exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment. Models S3 and S4 show minimal to no significant autocorrelations, suggesting more random recruitment patterns. Model S6 shows moderate autocorrelations, indicating some predictability based on recent past values. These findings help understand the temporal dynamics of krill recruitment and inform future modeling efforts. (fig autocorrelation)
## Kobe (status)

@@

@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1. @Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula. Although these models had convergence issues, they represented significant advances in this type of framework. Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings. Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Ossilation Southern (ENSO).

In that sense, the work carried out here differs substantially from the previously written integrated models since it incorporates aspects that have to do with ecosystem indicators such as environmental variables and predators. On the other hand, it considers the spatial component within the modeling scheme implicitly and finally evaluates the impact of the different sets of life history parameters that have been used and those calculated by Mardones et al (in press)

From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean. The intention to incorporate elements of predation in a stock is necessary to improve the modelling, because did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000]. There is many studies that confirm the influence of predation on grill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022]. For this, we consider penguin columns as one of the most powerful elements of natural predation on krill because have direct correlations between this type of taxonomic group and krill [@Kruger2021].

@@

 By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model. This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.

### 2024-05-30 index.Rmd
commit 31 may

---
title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
author: "Mardones, M; Cárdenas, C., Krüger, L., Santa Cruz, F."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
```

```{r message=FALSE, warning=FALSE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # whitout S-R
dir4<-here("s4") # 
dir5<-here("s5") # whitout S-R
dir6<-here("s6") # whitout S-R
dir7<-here("s7") # whitout S-R
```

@@

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata and Predator and Env data |
|    s3     | Same "s2" Without S-R relation         |
|    s4     | Same "s2" Ricker S-R relation estimated          |
|    s5     | Same "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | Same "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | Same "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |

@@

```{r eval=FALSE, message=F, include=FALSE}
# Lista de directorios para correro tordos juntos
directorios <- c("s01", 
                 "s1", 
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar

# Bucle para ejecutar el código en cada directorio
for (dir in directorios) {
  r4ss::run(
    dir = dir,
    exe = "ss_osx",
    skipfinished = FALSE,
    show_in_console = TRUE
  )
}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
#system('./ss_osx') # to mac
r4ss::run(
  dir = dir2,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```

@@

#s7
base.model7 <- SS_output(dir=dir7,
                        covar=T,
                        forecast=T)
```

```{r eval=FALSE, message=F, include=FALSE}
SS_plots(base.model2, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)
```

@@

```{r}
SSplotTimeseries(base.model01, subplot = 7)
```
Salida de las biomasas con todas las flotas

@@

```{r eval=FALSE}
SSplotRunstest(base.model01, 
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2, 
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
```

@@

```{r}
mvln = SSdeltaMVLN(base.model01,
                   Fref = "MSY", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model2,
                   Fref = "MSY", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)
```

@@

```{r}
like01 <- base.model01$likelihoods_used
like01$model <- rep("s01", nrow(like01))
like01 <- rownames_to_column(like01, var = "Description")
like1 <- base.model1$likelihoods_used
like1$model <- rep("s1", nrow(like1))
like1 <- rownames_to_column(like1, var = "Description")
like2 <- base.model1$likelihoods_used
like2$model <- rep("s2", nrow(like2))
like2 <- rownames_to_column(like2, var = "Description")
like3 <- base.model3$likelihoods_used
like3$model <- rep("s3", nrow(like3))
like3 <- rownames_to_column(like3, var = "Description")
like4 <- base.model4$likelihoods_used
like4$model <- rep("s4", nrow(like4))
like4 <- rownames_to_column(like4, var = "Description")
like5 <- base.model5$likelihoods_used
like5$model <- rep("s5", nrow(like5))
like5 <- rownames_to_column(like5, var = "Description")
like6 <- base.model6$likelihoods_used
like6$model <- rep("s6", nrow(like6))
like6 <- rownames_to_column(like6, var = "Description")
like7 <- base.model7$likelihoods_used
like7$model <- rep("s7", nrow(like7))
like7 <- rownames_to_column(like7, var = "Description")
totalike <- rbind(like01,
                  like1,
                   like2,
                   like3,
                   like4,
                   like5,
                   like6,
                   like7)

@@

```{r}
ggplot(totalike)+
  geom_bar( aes(y = reorder(Description, values), x =values),
            stat = "identity", position = "dodge") +
  theme_few() +
  facet_wrap(.~model, 
             ncol=4)+
  labs(y="",
       x="Likelihood")+
  xlim(0, 1000)+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1)) 
```

@@

```{r echo=FALSE}
#PLOT labels, name of each model run
legend.labels <- c( "s01",
                  "s1",
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7")

#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL,
                       dirvec = c(dir01,
                                  dir1,
                                  dir2,
                                  dir3,
                                  dir4,
                                  dir5,
                                  dir6,
                                  dir7),
                       getcovar = F)

@@

SSplotComparisons(summaryoutput,
                  legendlabels = c("s01",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 plotdir = "~/Figs")

comp <- SSplotComparisons(summaryoutput,
                  legendlabels = c("s01",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                 subplots = c(2, 16),
                 plotdir = "~/Figs")
comp
```

@@

@@

>>>>

---
title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
author: "Mardones, M; Cárdenas, C., Krüger, L., Santa Cruz, F."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
```

```{r message=FALSE, warning=FALSE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # whitout S-R
dir4<-here("s4") # 
dir5<-here("s5") # whitout S-R
dir6<-here("s6") # whitout S-R
dir7<-here("s7") # whitout S-R
dir8<-here("s8") # S2 set parametres EMM-2024/23 (Mardones)
```

@@

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3     | "s2" without S-R relation         |
|    s4     | "s2" Ricker S-R relation estimated          |
|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |

@@

```{r eval=FALSE, message=F, include=FALSE}
# Lista de directorios para correro tordos juntos
directorios <- c("s01", 
                 "s1", 
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7",
                 "s8")  # Agrega aquí todos los nombres de las carpetas que deseas procesar

# Bucle para ejecutar el código en cada directorio
for (dir in directorios) {
  r4ss::run(
    dir = dir,
    exe = "ss_osx",
    skipfinished = FALSE,
    show_in_console = TRUE
  )
}
```

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
#system('./ss_osx') # to mac
r4ss::run(
  dir = dir01,
  exe = "ss_osx",
  skipfinished = FALSE, # TRUE will skip running if Report.sso present
  show_in_console = TRUE # change to true to watch the output go past
)
```

@@

#s7
base.model7 <- SS_output(dir=dir7,
                        covar=T,
                        forecast=T)

#s7
base.model8 <- SS_output(dir=dir8,
                        covar=T,
                        forecast=T)
```

```{r eval=FALSE, message=F, include=FALSE}
SS_plots(base.model3, 
         uncertainty = TRUE,
         datplot = T, 
         png=T, 
         aalresids = F,
         btarg=0.75, 
         minbthresh=0.2, 
         forecast=T)
```

@@

```{r}
SSplotTimeseries(base.model2, subplot = 7)
```
Salida de las biomasas con todas las flotas

@@

```{r eval=FALSE}
SSplotRunstest(base.model8, 
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2, 
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model01,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model1,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model3,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model4,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model5,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model6,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model7,
               subplots = "len",
               add=T)
SSplotJABBAres(base.model7,
               subplots = "len",
               add=T)

SSplotMnwt(base.model01, subplots = 2)
```

@@

```{r}
mvln = SSdeltaMVLN(base.model8,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=1000)

mvln = SSdeltaMVLN(base.model2,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=1000)
```

@@

```{r}
like01 <- base.model01$likelihoods_used
like01$model <- rep("s01", nrow(like01))
like01 <- rownames_to_column(like01, var = "Description")
like1 <- base.model1$likelihoods_used
like1$model <- rep("s1", nrow(like1))
like1 <- rownames_to_column(like1, var = "Description")
like2 <- base.model1$likelihoods_used
like2$model <- rep("s2", nrow(like2))
like2 <- rownames_to_column(like2, var = "Description")
like3 <- base.model3$likelihoods_used
like3$model <- rep("s3", nrow(like3))
like3 <- rownames_to_column(like3, var = "Description")
like4 <- base.model4$likelihoods_used
like4$model <- rep("s4", nrow(like4))
like4 <- rownames_to_column(like4, var = "Description")
like5 <- base.model5$likelihoods_used
like5$model <- rep("s5", nrow(like5))
like5 <- rownames_to_column(like5, var = "Description")
like6 <- base.model6$likelihoods_used
like6$model <- rep("s6", nrow(like6))
like6 <- rownames_to_column(like6, var = "Description")
like7 <- base.model7$likelihoods_used
like7$model <- rep("s7", nrow(like7))
like7 <- rownames_to_column(like7, var = "Description")
like8 <- base.model8$likelihoods_used
like8$model <- rep("s8", nrow(like8))
like8 <- rownames_to_column(like8, var = "Description")
totalike <- rbind(like01,
                  like1,
                   like2,
                   like3,
                   like4,
                   like5,
                   like6,
                   like7,
                  like8)

@@

```{r}
ggplot(totalike)+
  geom_bar( aes(y = reorder(Description, values), x =values),
            stat = "identity", position = "dodge") +
  theme_few() +
  facet_wrap(.~model, 
             ncol=3)+
  labs(y="",
       x="Likelihood")+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1)) 
```

@@

```{r echo=FALSE}
#PLOT labels, name of each model run
legend.labels <- c( "s01",
                  "s1",
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7",
                 "s8")

#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL,
                       dirvec = c(dir01,
                                  dir1,
                                  dir2,
                                  dir3,
                                  dir4,
                                  dir5,
                                  dir6,
                                  dir7,
                                  dir8),
                       getcovar = F)

@@

SSplotComparisons(summaryoutput,
                  legendlabels = c("No Spatial",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7",
                 "s8"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 plotdir = "~/Figs")

@@

```
comparision between select models

```{r echo=FALSE}
#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL,
                       dirvec = c(
                                  dir8,

                                  dir2,
                                  dir3),
                       getcovar = F)

SSplotComparisons(summaryoutput,
                  legendlabels = c(
                                   "s8",

                 "s2",
                 "s3"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 plotdir = "~/Figs")

comp2_8 <- SSplotComparisons(summaryoutput,
                  legendlabels = c(
                                   "s8",

                 "s2",
                 "s3"),
                 subplots = c(2, 16),
                 plotdir = "~/Figs")
comp2_8
```
## Comparsion in sd long term time series

```{r}
sblt01 <- base.model01$timeseries$Bio_all 
sblt1 <- base.model1$timeseries$Bio_all 
sblt2 <- base.model2$timeseries$Bio_all 
sblt3 <- base.model3$timeseries$Bio_all 
sblt4 <- base.model4$timeseries$Bio_all 
sblt5 <- base.model5$timeseries$Bio_all 
sblt6 <- base.model6$timeseries$Bio_all 
sblt7 <- base.model7$timeseries$Bio_all 
sblt8 <- base.model8$timeseries$Bio_all
errt <- as.data.frame(cbind(sblt01,
              sblt1,
              sblt2,
              sblt3,
              sblt4,
              sblt5,
              sblt6,
              sblt7,
              sblt8))
err_long <- errt %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value")

ggplot(err_long, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.3) +
  labs(title = "",
       x = "",
       y = "Spawning B.(mt)") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1))
base.model2$

@@

## Autocorrelation in Recruit and Biomas

```{r}
recs01 <- base.model01$recruit$pred_recr
recs1 <- base.model1$recruit$pred_recr
recs2 <- base.model2$recruit$pred_recr
recs3 <- base.model3$recruit$pred_recr
recs4 <- base.model4$recruit$pred_recr
recs5 <- base.model5$recruit$pred_recr
recs6 <- base.model6$recruit$pred_recr
recs7 <- base.model7$recruit$pred_recr
recs8 <- base.model8$recruit$pred_recr

p01 <- ggAcf(base.model01$recruit$pred_recr) +
  ggtitle("S01") +
  theme_few()
p1 <- ggAcf(base.model1$recruit$pred_recr) +
  ggtitle("S1") +
  theme_few()
p2 <- ggAcf(base.model2$recruit$pred_recr) +
  ggtitle("S2") +
  theme_few()
p3 <- ggAcf(base.model3$recruit$pred_recr) +
  ggtitle("S3") +
  theme_few()
p4 <- ggAcf(base.model4$recruit$pred_recr) +
  ggtitle("S4") +
  theme_few()
p5 <- ggAcf(base.model5$recruit$pred_recr) +
  ggtitle("S5") +
  theme_few()
p6 <- ggAcf(base.model6$recruit$pred_recr) +
  ggtitle("S6") +
  theme_few()
p7 <- ggAcf(base.model7$recruit$pred_recr) +
  ggtitle("S6") +
  theme_few()
p8 <- ggAcf(base.model8$recruit$pred_recr) +
  ggtitle("S6") +
  theme_few()

ggarrange(p01,p1,
          p2,
          p3,
          p4,
          p5,
          p6,
          p7,
          p8, ncol=3, nrow = 3,
          common.legend = T)
```
